---
title: "Networks - day 1"
author: "CB"
date: "2023-02-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Manipulation and characterization of the networks

We first need to install the `igraph` library:

```{r}
#install.packages('igraph')
library(igraph)
```

After, we can create a first graph and try to manipulate it.

> Exercise: create an adjacency matrix corresponding to a undirected network where the first node is connected to the 4 other nodes, that are not connected togather.

```{r}
A = cbind(c(0,1,1,1,1),
          c(1,0,0,0,0),
          c(1,0,0,0,0),
          c(1,0,0,0,0),
          c(1,0,0,0,0))
```

This adjacency matrix can be used within the `igraph` environment for plots or computations.

```{r}
g = graph.adjacency(A)
g
```

```{r}
plot(g)
```

If we prefer to display a undirected network, we have make this call:

```{r}
g = graph.adjacency(A, mode='undirected')
plot(g)
```

> Exercise= modify the adjacendy matrix to make a connection between the nodes 3 and 5, and plot it.

```{r}
A = cbind(c(0,1,1,1,1),
          c(1,0,0,0,0),
          c(1,0,0,0,1),
          c(1,0,0,0,0),
          c(1,0,1,0,0))
g = graph.adjacency(A, mode='undirected')
plot(g)
```

We can also use the `igraph` library to compute the degree of the nodes.

```{r}
degree(g)
```

Another statistic we can compute for graphs is the number of k-cliques.

> Definition: a k-clique is a set of k nodes that are all connected togather.

For instance, in the above network, there is one 3-clique and 5 2-cliques.

```{r}
cliques(g)
```

> Exercice: write a small piece of code allowing to plot the signature of the graph by counting the number of k-cliques, k=1,...,n

```{r}
signature <- function(g, max = 10){
   cl = cliques(g)
   s = rep(0,max)
   for (i in 1:length(cl)){
      l = length(cl[[i]])
      s[l] = s[l] + 1
   }
   barplot(s)
   return(s)
}
signature(g)
```





