\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}

% Importing settings from setup.sty
\usepackage{setup}
\usepackage{booktabs}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{glossaries}
% \makenoidxglossaries
\newcommand{\prox}{\operatorname{prox}}

% \newacronym{bmi}{BMI}{Body Mass Index}


% \pagenumbering{roman}
\begin{document}

% Inserting title page
\import{./}{title}

\pagenumbering{gobble}
\tableofcontents
% \listoffigures
% \listoftables



\newgeometry{
  left=25mm,
  right=25mm,
  top=25mm,
  bottom=25mm}
\pagenumbering{arabic}

\section{Exercise 1: Soft-thresholding}
\label{sec:soft-thresholding}
The proximal operator of $\tau f$ is defined as:
\begin{equation*}
  \prox_{\tau f}(x) = \arg\min_{u \in \R} \frac{1}{2\tau}\left(u-x \right)^2 + f(u)
\end{equation*}
which we can apply to the $\ell_1$ norm to get:
\begin{equation*}
  \prox_{\tau |\cdot|}(x) = \arg\min_{u \in \R} \frac{1}{2\tau}\left(u-x \right)^2 + |u|
\end{equation*}
Let $h(u)$ be given by:
$$
  h(u) := \frac{1}{2\tau}\left(u-x \right)^2 + |u|
$$
The optimality condition states that given that $h$ is is proper, we have:
\begin{equation*}
  0 \in \partial h(u^*) \iff u^* \in \arg \min_{u \in \R} h(u)
\end{equation*}
then we have:
\begin{align*}
  \frac{\partial}{\partial u} h(u)
   & = \frac{\partial}{\partial u}  \left[\frac{1}{2\tau}\left(u-x \right)^2 + |u| \right] \\
   & =
  \begin{cases}
    \frac{1}{\tau}(u-x) - 1,  & u < 0 \\
    \frac{1}{\tau} (u-x) + 1, & u > 0 \\
  \end{cases}
\end{align*}
We set the derivative to zero to find the critical points of $h$. We have three cases to consider: $u > 0$, $u < 0$ and $u = 0$.
\paragraph{Case $u > 0$.}
\begin{align*}
           &
  \frac{\partial}{\partial u}  h(u^*) = 0 \\
  \implies &
  \frac{1}{\tau} (u^*-x) + 1 = 0          \\
  \implies &
  u^* = x - \tau                          \\
\end{align*}

\paragraph{Case $u < 0$.}
\begin{align*}
           &
  \frac{\partial}{\partial u}  h(u^*) = 0 \\
  \implies &
  \frac{1}{\tau} (u^*-x) - 1 = 0          \\
  \implies &
  u^* = x + \tau                          \\
\end{align*}

\paragraph{Case $u = 0$.}
In this case, we cannot compute the derivative as the function is non-differentiable in $u = 0$. We have however that the subdifferential of $h$ is given by:
$$
  \partial h(u) = [-1, 1]
$$
In particular, $0 \in \partial h(u)$, so we can apply the optimality condition. Therefore, the proximal operator is given by:
\begin{align*}
  \prox_{\tau |\cdot|}(x)
   & = \begin{cases}
    x - \tau, & u > 0            \\
    x + \tau, & u < 0            \\
    0,        & \text{otherwise} \\
  \end{cases}  \\
   & = \begin{cases}
    x - \tau, & x - \tau > 0     \\
    x + \tau, & x + \tau < 0     \\
    0,        & \text{otherwise} \\
  \end{cases}  \\
   & = \begin{cases}
    x - \tau, & x > \tau         \\
    x + \tau, & x < - \tau       \\
    0,        & \text{otherwise} \\
  \end{cases}  \\
   & = \begin{cases}
    x - \tau, & x > \tau      \\
    x + \tau, & x < -\tau     \\
    0,        & |x| \leq \tau \\
  \end{cases} \\
\end{align*}
\\We plot this proximal operator in the companion notebook.

\section{Exercise 2: Hard-thresholding}
We define $f$ as the $\ell_0$ norm:
\begin{equation*}
  f(x) = |x|_0 = \begin{cases}
    0, & x = 0    \\
    1, & x \neq 0
  \end{cases}
\end{equation*}
The proximal operator of $\tau f$ is defined as:
\begin{equation*}
  \prox_{\tau |\cdot|_0}(x) = \arg\min_{u \in \R} \frac{1}{2\tau}\left(u-x \right)^2 + |u|_0
\end{equation*}
% \begin{equation*}
%   \prox_{\tau |\cdot|_0}(x) = \begin{cases}
%     0, & x = 0    \\
%     x, & x \neq 0
%   \end{cases}
% \end{equation*}
Let $h(u)$ be given by:
\begin{align*}
  h(u)
   & := \frac{1}{2\tau}\left(u-x \right)^2 + |u|_0 \\
   & =
  \begin{cases}
    \frac{1}{2\tau} \left(u-x \right)^2 + 1,  & u \neq 0 \\
    \frac{1}{2\tau} \left( 0-x \right)^2 + 0, & u = 0    \\
  \end{cases}                       \\
   & =
  \begin{cases}
    \frac{1}{2\tau} \left(u-x \right)^2 + 1, & u \neq 0 \\
    \frac{x^2}{2\tau},                       & u = 0    \\
  \end{cases}                       \\
\end{align*}
We have two cases to consider: $u \neq 0$ and $u = 0$.
\paragraph{Case $u \neq 0$.}
In this case, $h$ is differentiable. We will compute its derivative and set it to zero to find the critical points of $h$. The derivative of $h$ is given by:
\begin{align*}
  \frac{\partial}{\partial u} h(x)
   & = \frac{1}{\tau} (u-x)
\end{align*}
We now set this derivative to zero to find the critical points of $h$:
\begin{align*}
           &
  \frac{\partial}{\partial u}  h(u^*) = 0 \\
  \implies &
  \frac{1}{\tau} (u^*-x) = 0              \\
  \implies &
  u^* = x                                 \\
\end{align*}
Therefore, we have:
\begin{align*}
  h(u^*)
   & = \frac{1}{2\tau}\left(u^* - x \right)^2 + |u^*|_0 \\
   & = \frac{1}{2\tau}\left(x - x \right)^2 + 1         \\
   & = 1                                                \\
\end{align*}
\paragraph{Case $u = 0$.}
In this case, we cannot compute the derivative as the function is non-differentiable in $u = 0$. We have that $h$ is given by:
\begin{equation*}
  h(u) = \frac{x^2}{2\tau}
\end{equation*}
Now, the question is whether it is better to have $h(u) = \frac{x^2}{2\tau}$ or $h(u) = 1$. We compare the two quantities:
\begin{align*}
  \frac{x^2}{2\tau} < 1
  \implies &
  x^2 < 2\tau                         \\
  \implies &
  x \in [-\sqrt{2\tau}, \sqrt{2\tau}] \\
\end{align*}
So in order to minimize $h$, we prefer having $h(u) = \frac{x^2}{2\tau}$ as long as $x \in [-\sqrt{2\tau}, \sqrt{2\tau}]$ and $h(u) = 1$ otherwise. We need to choose $u$ appropriately, which means that the proximal operator of $\tau |\cdot|_0$ is given by:
\begin{equation*}
  \prox_{\tau |\cdot|_0}(x) = \begin{cases}
    0, & x \in [-\sqrt{2\tau}, \sqrt{2\tau}] \\
    x, & \text{otherwise}                    \\
  \end{cases}
\end{equation*}
\\We plot this proximal operator in the companion notebook.


\section{Exercise 3: Non-negativity constraints}
\subsection{Part 1}
\label{sec:part1}
Let $\R_+^n$ be the set of vectors with non-negative entries. We define the indicator function of $\R_+^n$ as:
\begin{equation*}
  \delta_{\R_+^n} (x) =
  \begin{cases}
    0,      & x \in \R_+^n     \\
    \infty, & \text{otherwise}
  \end{cases}
\end{equation*}
Therefore the proximal operator of $\delta_{\R_+^n}$ is given by:
\begin{equation*}
  \prox_{\delta_{\R_+^n}}(x) = \arg\min_{u \in \R^n} \frac{1}{2}\|u-x\|^2 + \delta_{\R_+^n} (u)
\end{equation*}
We define $h(u)$ as:
\begin{equation*}
  h(u) = \frac{1}{2}\|u-x\|^2 + \delta_{\R_+^n} (u)
\end{equation*}
We understand from the definition of the indicator function that no component of $u$ can be negative, otherwise $h$ would be infinite. \\
We have two cases to consider: $u \in \R_+^n$ and $u \notin \R_+^n$.
\paragraph{Case $u \in \R_+^n$.}
In this case, $h$ is differentiable. We will compute its derivative and set it to zero to find the critical points of $h$. The derivative of $h$ is given by:
\begin{align*}
  \nabla h(u^*) = 0
  \implies &
  u^* - x = 0 \\
  \implies &
  u^* = x
\end{align*}
\paragraph{Case $u \notin \R_+^n$.}
In this case, as mentioned above, $h$ is infinite. We cannot choose any component $u_i$ of $u$ to be negative, otherwise $h$ would be infinite because of the indicator function. However, we still want to choose $u$ to be as close as possible to $x$, in order to minimize the $\|u-x\|^2$ component. We proceed in a component-wise fashion, thanks to the separability of $h$. For all $i$ such that $x_i < 0$, we can therefore choose $u_i = 0$, which is the point in $\R_+$ which minimizes the distance to $x_i$. \\
We can summarize the proximal operator of $\delta_{\R_+^n}$ as:
\begin{equation*}
  \forall i=1, \ldots, n,\ \left\{\prox_{\delta_{\R_+^n}}(x)\right\}_i = \max(0, x_i)
\end{equation*}
\subsection{Part 2}
We now compute the proximal operator of $\tau |\cdot|_1 + \delta_{\R_+^n} (\cdot)$, which is given by:
\begin{equation*}
  \prox_{\tau |\cdot|_1 + \delta_{\R_+^n} (\cdot)}(x) = \arg\min_{u \in \R_+^n} \frac{1}{2\tau}\|u-x\|^2 + |u|_1 + \delta_{\R_+^n} (u)
\end{equation*}
We define $h(u)$ as:
\begin{equation*}
  h(u) = \frac{1}{2\tau}\|u-x\|^2 + |u|_1 + \delta_{\R_+^n} (u)
\end{equation*}
We note that $h$ is separable. We can therefore apply the proximal operator of $\tau |\cdot|_1$ to each component of $u$. Moreover, we can use the same argument as in section \ref{sec:part1} regarding the non-negativity constraint. Indeed once again, as long as one of the component of $u$ is negative, $h$ is infinite, so minimizing $h$ means that all components of $u$ must be non-negative. For the rest, the minimizer of $\frac{1}{2\tau}\|u-x\|^2 + |u|_1$ is by definition equal to $\prox_{\tau |\cdot|_1}(x)$, that is:
\begin{equation*}
  \arg\min_{u \in \R^n} \frac{1}{2\tau}\|u-x\|^2 + |u|_1 =: \prox_{\tau |\cdot|_1}(x)
\end{equation*}
and we computed this proximal operator in section \ref{sec:soft-thresholding}. \\
Now, the indicator forces all components of $u$ to be non-negative (this is the same argument as in section \ref{sec:part1} really). As a result, we can summarize the proximal operator of $\tau |\cdot|_1 + \delta_{\R_+^n} (\cdot)$ as:
\begin{equation*}
  \forall i = 1, \ldots, n, \  \left(\prox_{\tau |\cdot|_1 + \delta_{\R_+^n} (\cdot)}(x)\right)_i = \max\left(\left(\prox_{\tau |\cdot|_1}(x)\right)_i, 0\right)
\end{equation*}
where the maximum is taken component-wise.



\section{Exercise 4}
Let us define $f$ as the elastic net functional, that is:
\begin{equation*}
  f(x) = \|x\|_1 + \frac{\lambda}{2}\|x\|^2
\end{equation*}
We want ot compute $\prox_{f}(x)$. \\
We define $g$ as:
\begin{equation*}
  g(x) = \| x \|_1
\end{equation*}
therefore, we have that:
\begin{equation*}
  f(x) = g(x) + \frac{\lambda}{2}\|x\|^2
\end{equation*}
By proposition 3 with $c = \lambda$, we have that:
\begin{align*}
  \prox_{f}(x) = \prox_{\frac{g}{\lambda + 1}} \left( \frac{x}{\lambda + 1}\right)
\end{align*}
Now, using proposition 2 with the variable from the proposition $\lambda_{\text{prop}}$ being such that $\lambda_{\text{prop}} = \lambda + 1$, we have that:
\begin{align*}
           &
  (\lambda + 1) \prox_{\frac{g}{\lambda + 1}} \left( \frac{x}{\lambda + 1}\right) = \prox_{g} \left( x \right)         \\
  \implies &
  \prox_{\frac{g}{\lambda + 1}} \left( \frac{x}{\lambda + 1}\right) = \frac{1}{\lambda + 1} \prox_{g} \left( x \right) \\
  \implies &
  \prox_{f} \left( x \right) = \frac{1}{\lambda + 1} \prox_{g} \left( x \right)                                        \\
  \implies &
  \prox_{f} \left( x \right) = \frac{1}{\lambda + 1} \prox_{| \cdot |_1} \left( x \right)                              \\
\end{align*}
where we know section \ref{sec:soft-thresholding} the value of the proximal operator of $| \cdot |_1$.


% \clearpage
% \printnoidxglossaries

\end{document}