---
title: "MBSL-lesson9"
author: "CB"
date: "2022-11-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Subspace clustering

We need to install the `MBCbook` package for accessing some high-dimensional data and the `HDclassif` package for HDDC/MPPCA:

```{r}
#install.packages(c('MBCbook','HDclassif'))
library(MBCbook)
library(HDclassif)
```

Thanks to the first package, we have access to the `usps358` data.

```{r}
data(usps358)
?usps358
```

Since all data are vectorized images, it is naturally possible to revert the process and visualize the images:

```{r}
X = usps358[,-1]
Y = usps358$cls
image(matrix(as.double(X[123,]),nrow=16))
```

> Exercise: try to cluster these data set with kmeans and Mclust, and compare the results with the actual labels.

```{r}
out.kmeans = kmeans(X,centers = 3,nstart = 10)
   
# Compare with the true labels
table(Y,out.kmeans$cluster)

# Let's try to visualize the cluster means
par(mfrow=c(1,3))
image(matrix(as.double(out.kmeans$centers[1,]),nrow=16))
image(matrix(as.double(out.kmeans$centers[2,]),nrow=16))
image(matrix(as.double(out.kmeans$centers[3,]),nrow=16))
```


```{r}
out.mclust = Mclust(X,G=3,initialization=list(subset = sample(nrow(X),300)))

# Compare with the true labels
table(Y,out.mclust$classification)

# Let's try to visualize the cluster means
par(mfrow=c(1,3))
image(matrix(as.double(out.mclust$parameters$mean[,1]),nrow=16))
image(matrix(as.double(out.mclust$parameters$mean[,2]),nrow=16))
image(matrix(as.double(out.mclust$parameters$mean[,3]),nrow=16))
```
Finally, let's test HDDC/MPPCA:

```{r}
#out.hddc = hddc(X,K = 3,model = "AkjBkQkd") # MPPCA
out.hddc = hddc(X,K = 3,model = "AkjBkQkdk")

# Compare with the true labels
table(Y,out.hddc$class)

# Let's try to visualize the cluster means
par(mfrow=c(1,3))
image(matrix(as.double(out.hddc$mu[1,]),nrow=16))
image(matrix(as.double(out.hddc$mu[2,]),nrow=16))
image(matrix(as.double(out.hddc$mu[3,]),nrow=16))
```

It is interesting to look at the estimated dimensions for the cluster subspaces:

```{r}
out.hddc$d
```









